Non-word repetition in Yélî Dnye

### 2021-01-20 manuscript.Rmd
started discussion

For consonants, we inspected complex (`r print.phone.list(complexCs.ipa)`) versus simpler ones (`r print.phone.list(simplexCs.ipa)`), using the same logic: We looked at correct phone repetition, substitution with a change in complexity category, or a change within the same complexity category.^[Note that the substitutions included phones that are not native to Yélî Dnye but do occur in English (e.g., [tʃ]); as these data come from careful transcriptions by a native Yélî Dnye speaker who is very fluent in English, we take these segmental substitutions faithfully as an indication that several of our participants have mastered production of some English phones, possibly produced within whole English word forms.] 
The complex consonants in our stimuli occur in ~`r mean(segments$pc[segments$type %in% c("complex")],na.rm=T)`% of languages' phonologies (range `r min(segments$pc[segments$type %in% c("complex")],na.rm=T)`% to `r max(segments$pc[segments$type %in% c("complex")],na.rm=T)`%); 
whereas simple consonants in our stimuli occur in ~`r mean(segments$pc[segments$type %in% c("simple")],na.rm=T)`% of languages' phonologies (range `r min(segments$pc[segments$type %in% c("simple")],na.rm=T)`% to `r max(segments$pc[segments$type %in% c("simple")],na.rm=T)`%). 
Again these groups of sounds differ in their frequency within the language. Their type frequencies in Yélî Dnye are: complex consonants ~`r mean(segments$freq_corpus_types[segments$type %in% c("complex")],na.rm=T)*10`‰  (range `r min(segments$freq_corpus_types[segments$type %in% c("complex")],na.rm=T)*10`‰ to `r max(segments$freq_corpus_types[segments$type %in% c("complex")],na.rm=T)*10`‰) versus simple consonants ~`r mean(segments$freq_corpus_types[segments$type %in% c("simple")],na.rm=T)*10`‰ (range `r min(segments$freq_corpus_types[segments$type %in% c("simple")],na.rm=T)*10`‰ to `r max(segments$freq_corpus_types[segments$type %in% c("simple")],na.rm=T)*10`‰).

@@

Next, we inspected whether NWR scores varied as a function of word length (Table \@ref(tab:tablength)). Participants scored much lower on monosyllables than on non-words of other lengths. This is likely due to the fact that the majority of monosyllables were designed to include sounds that are rare in the world's languages, which may indicate that they are hard to produce or perceive. Setting monosyllables aside, we observed the typical pattern of lower scores for longer items only for the whole-item scoring, and even there differences were rather small. In a generalized binomial mixed model excluding monosyllables, we included `r summary(len_lm)$devcomp$dims["N"]` observations, from `r summary(len_lm)$ngrps["file"]` children producing, in any given trial, one of `r summary(len_lm)$ngrps["target"]` potential target words. The analysis revealed a positive effect of age (ß = `r mycoeff["age.rounded_s","Estimate"]`, SE ß = `r mycoeff["age.rounded_s","Std. Error"]`, p `r pval.display(mycoeff["age.rounded_s","Pr(>|z|)"])`) and a negative but non-significant estimate for target length in number of syllables (ß = `r mycoeff["type","Estimate"]`, SE ß = `r mycoeff["type","Std. Error"]`, p `r pval.display(mycoeff["type","Pr(>|z|)"])`).

@@

```{r Fig2-scores-by-age, ig.pos = "T!", fig.width=5, fig.height=5, fig.cap="NWR whole-item scores for individual participants as a function of age and sex (purple = boys, orange = girls)."}

plot(overall$Word~overall$Age,pch=4,cex=2,lty=3,ylab="NWR scores (% words)",xlab="Age (years)",col=ifelse(overall$Sex=="B","darkmagenta","chocolate1"))

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure \@ref(fig:Fig2-scores-by-age), there was a greater deal of variance across the tested age range, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (`r corage$statistic`) = `r corage$estimate`, p `r pval.display(corage$p.value)`). In contrast, there was no clear association between NWR scores and sex (Welch t (`r tsex$parameter`) = `r tsex$statistic`, p `r pval.display(tsex$p.value)`), birth order  (data missing for `r sum(is.na(demo$birthOrder[demo$included=="yes"]))` children, rho = (`r round(corbo$statistic,1)`) = `r corbo$estimate`, p `r pval.display(corbo$p.value)`), or maternal education  (rho ( `r round(corme$statistic,1)`) = `r corme$estimate`, p `r pval.display(corme$p.value)`).

@@

We used non-word repetition to investigate phonological development in a language with a large phonological inventory (including some typologically rare segments). While the study, in itself, advances NWR research by demonstrating the successful application of this method in a rural, non-industrialized population [see also @cristia2020infant], the results also inform the larger body of NWR work regarding the influence of stimulus length, segmental frequency, age, and developmental context, where previous findings have often been mixed.

### Item complexity

We investigated the effect of item complexity on NWR scores by varying both the number of syllables in the item and its average segmental frequency. Based on previous work, we had predicted that children would have higher NWR scores for shorter items. That said, previous work has shown both very small [@piazzalunga2019articulatory] and very large [@cristia2020infant; @jabere2018xperiment] effects of stimulus length and, further, the Yélî Dnye dictionary suggests that mono- and bi-syllabic words are nearly equally frequent in the current language, with trisyllabic and longer words making up a non-trivial 10% of the remaining words. Compare this to, for example, English, which is substantially more skewed toward monosyllabic word forms **M2A: Alex I'm going off your note here ("Prediction for Yélî made before seeing the data: The length distribution in Yélî words is more balanced than that in English, and thus the score decline for poly- versus mono-syllables may be less pronounced than that for English.""). I don't have a reference for this, can you please finish the thought or nix this bit?**. Setting aside our monosyllabic stimuli, which all contained typologically infrequent segments, we can examine effects of item length among the remaining stimuli, which range between 2 and 4 syllables long. While indeed NWR scores were overall lower for longer items (e.g., see Figure \@ref(fig:prevlit-fig)), the effect of item length was not significant in a statistical model that additionally accounted for age and random effects of item and participant. In light of mixed prior results of item length, we propose two possible (and non-mutually exclusive) explanations for this minimal impact of item length. First, further extensions of this type of analysis in more populations may reveal that, in general (and cross-linguistically), item length effects are variable between languages, potentially reflecting the distribution of word lengths in the ambient language and other (morpho-)phonological tendencies in the lexicon. Second, above and beyond these language-specific effects, the general impact of item length on NWR score may be relatively small, as shown in Piazzalunga et al.'s [-@piazzalunga2019articulatory] study on Italian and as borne out in the current dataset once controlling for other factors.

Our monosyllabic items included typologically rare segments so that we could test whether lower average segmental frequency is associated with lower NWR scores. Typologically common sounds are associated with higher performance on a handful of other tasks (**REFS -- M2A: Alex, I added this based on your note, where it sounded like you had some particular studies in mind?**) though to our knowledge this has not yet been tested with non-word repetition. Regarding Yélî Dnye in particular, the phonemic inventory is both large and acoustically packed, in addition to containing several typologically infrequent (or unique) contrasts. We therefore expected to see that, while NWR scores would be lower for stimuli with lower average frequency, this effect would be relatively weak because the ambient language puts pressure on Yélî children to distinguish (perceptually and articulatorily) fine-grained phonetic differences in order to successfully communicate with others. Indeed we found a robust effect of average segmental frequency on NWR performance: Even accounting for age and random effects of item and participant, we see that target words with more frequent segments were repeated correctly more often. This effect is large, with a magnitude more than twice the size of the effect of participant age. This significant effect remains even once also accounting for the frequencies of these segments in Yélî Dnye child-directed speech, which are correlated with their typological frequencies. In sum, typological frequency effects, which have been found in other measurements of phonological processing. appear to strongly affect NWR performance, and do not appear mitigated by language-specific pressure to make finer-grained differences earlier in development.

@@

>>>>

For consonants, we inspected complex (`r print.phone.list(complexCs.ipa)`) versus simpler ones (`r print.phone.list(simplexCs.ipa)`), using the same logic: We looked at correct phone repetition, substitution with a change in complexity category, or a change within the same complexity category.^[Note that the substitutions included phones that are not native to Yélî Dnye but do occur in English (e.g., [tʃ]). These data come from careful transcriptions by a native Yélî Dnye speaker who is very fluent in English. This result suggests that several of our participants have mastered production of some English phones, possibly produced within whole English word forms.] 
The complex consonants in our stimuli occur in ~`r mean(segments$pc[segments$type %in% c("complex")],na.rm=T)`% of languages' phonologies (range `r min(segments$pc[segments$type %in% c("complex")],na.rm=T)`% to `r max(segments$pc[segments$type %in% c("complex")],na.rm=T)`%); 
whereas simple consonants in our stimuli occur in ~`r mean(segments$pc[segments$type %in% c("simple")],na.rm=T)`% of languages' phonologies (range `r min(segments$pc[segments$type %in% c("simple")],na.rm=T)`% to `r max(segments$pc[segments$type %in% c("simple")],na.rm=T)`%). 
Again these groups of sounds differ in their frequency within the language. Their type frequencies in Yélî Dnye are: complex consonants ~`r mean(segments$freq_corpus_types[segments$type %in% c("complex")],na.rm=T)*10`‰  (range `r min(segments$freq_corpus_types[segments$type %in% c("complex")],na.rm=T)*10`‰ to `r max(segments$freq_corpus_types[segments$type %in% c("complex")],na.rm=T)*10`‰) versus simple consonants ~`r mean(segments$freq_corpus_types[segments$type %in% c("simple")],na.rm=T)*10`‰ (range `r min(segments$freq_corpus_types[segments$type %in% c("simple")],na.rm=T)*10`‰ to `r max(segments$freq_corpus_types[segments$type %in% c("simple")],na.rm=T)*10`‰).

@@

We next turned to our second research question by inspecting whether NWR scores varied as a function of word length (Table \@ref(tab:tablength)). In this section and all subsequent ones, we only look at first attempts, for the reasons discussed previously. Additionally, we noticed that participants scored much lower on monosyllables than on non-words of other lengths. This is likely due to the fact that the majority of monosyllables were designed to include sounds that are rare in the world's languages, which may be harder to produce or perceive, as suggested by our previous analyses of NWR scores as a function of cross-linguistic phone frequency and error patterns. Therefore, we set monosyllables aside for this analysis.

We observed the typical pattern of lower scores for longer items only for the whole-item scoring, and even there differences were rather small. In a generalized binomial mixed model excluding monosyllables, we included `r summary(len_lm)$devcomp$dims["N"]` observations, from `r summary(len_lm)$ngrps["file"]` children producing, in any given trial, one of `r summary(len_lm)$ngrps["target"]` (non-monosyllabic) potential target words. The analysis revealed a positive effect of age (ß = `r mycoeff["age.rounded_s","Estimate"]`, SE ß = `r mycoeff["age.rounded_s","Std. Error"]`, p `r pval.display(mycoeff["age.rounded_s","Pr(>|z|)"])`) and a negative but non-significant estimate for target length in number of syllables (ß = `r mycoeff["type","Estimate"]`, SE ß = `r mycoeff["type","Std. Error"]`, p `r pval.display(mycoeff["type","Pr(>|z|)"])`).

@@

```{r Fig2-scores-by-age, ig.pos = "T!", fig.width=5, fig.height=5, fig.cap="NWR whole-item scores for individual participants as a function of age and sex (purple crosses = boys, orange circles = girls)."}

plot(overall$Word~overall$Age,pch=ifelse(overall$Sex=="B",4,1),cex=2,lty=3,ylab="NWR scores (% words)",xlab="Age (years)",col=ifelse(overall$Sex=="B","darkmagenta","chocolate1"))

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals, as per our third research question. As shown in Figure \@ref(fig:Fig2-scores-by-age), there was a greater deal of variance across the tested age range, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (`r corage$statistic`) = `r corage$estimate`, p `r pval.display(corage$p.value)`). In contrast, there was no clear association between NWR scores and sex (Welch t (`r tsex$parameter`) = `r tsex$statistic`, p `r pval.display(tsex$p.value)`), birth order  (data missing for `r sum(is.na(demo$birthOrder[demo$included=="yes"]))` children, rho = (`r round(corbo$statistic,1)`) = `r corbo$estimate`, p `r pval.display(corbo$p.value)`), or maternal education  (rho ( `r round(corme$statistic,1)`) = `r corme$estimate`, p `r pval.display(corme$p.value)`).

@@

We used non-word repetition to investigate phonological development in a language with a large phonological inventory (including some typologically rare segments). We aimed to provide additional data on two questions already visitied in NWR work, namely the influence of stimulus length and individual variation, and one that has received less attention, regarding the possible relationship between phone frequency and NWR scores. An additional overarching goal was to revisit this task, which is very commonly used to document phonological development, particularly in children raised in urban settings, and who are learning IndoEuropean languages. We will thus consider our results not only in light of previous work, but also with attention to potential linguistic and population differences.

### Associations between NWR and cross-linguistic frequency

Arguably the most innovative aspect of our data relate to the inclusion of phones that are less commonly found across languages, and rarely used in NWR tasks. Our monosyllabic items included typologically rare segments so that we could test whether lower average segmental frequency is associated with lower NWR scores. Typologically common sounds are associated with higher performance on a handful of other tasks (**REFS -- M2A: Alex, I added this based on your note, where it sounded like you had some particular studies in mind?**) though to our knowledge this has not yet been tested with non-word repetition. Regarding Yélî Dnye in particular, the phonemic inventory is both large and acoustically packed, in addition to containing several typologically infrequent (or unique) contrasts. We therefore expected to see that, while NWR scores would be lower for stimuli with lower average frequency, this effect would be relatively weak because the ambient language puts pressure on Yélî children to distinguish (perceptually and articulatorily) fine-grained phonetic differences in order to successfully communicate with others. Indeed we found a robust effect of average segmental frequency on NWR performance: Even accounting for age and random effects of item and participant, we see that target words with more frequent segments were repeated correctly more often. This effect is large, with a magnitude more than twice the size of the effect of participant age. This significant effect remains even once also accounting for the frequencies of these segments in Yélî Dnye child-directed speech, which are correlated with their typological frequencies. In sum, typological frequency effects, which have been found in other measurements of phonological processing. appear to strongly affect NWR performance, and do not appear mitigated by language-specific pressure to make finer-grained differences earlier in development.

@@

### Item Length

We investigated the effect of item complexity on NWR scores by varying both the number of syllables in the item. Based on previous work, we had predicted that children would have higher NWR scores for shorter items. That said, previous work has shown both very small [@piazzalunga2019articulatory] and very large [@cristia2020infant; @jabere2018xperiment] effects of stimulus length and, further, the Yélî Dnye dictionary suggests that mono- and bi-syllabic words are nearly equally frequent in the current language, with trisyllabic and longer words making up a non-trivial 10% of the remaining words. Compare this to, for example, English, which is substantially more skewed toward monosyllabic word forms **M2A: Alex I'm going off your note here ("Prediction for Yélî made before seeing the data: The length distribution in Yélî words is more balanced than that in English, and thus the score decline for poly- versus mono-syllables may be less pronounced than that for English.""). I don't have a reference for this, can you please finish the thought or nix this bit?**. Setting aside our monosyllabic stimuli, which all contained typologically infrequent segments, we can examine effects of item length among the remaining stimuli, which range between 2 and 4 syllables long. While indeed NWR scores were overall lower for longer items (e.g., see Figure \@ref(fig:prevlit-fig)), the effect of item length was not significant in a statistical model that additionally accounted for age and random effects of item and participant. In light of mixed prior results of item length, we propose two possible (and non-mutually exclusive) explanations for this minimal impact of item length. First, further extensions of this type of analysis in more populations may reveal that, in general (and cross-linguistically), item length effects are variable between languages, potentially reflecting the distribution of word lengths in the ambient language and other (morpho-)phonological tendencies in the lexicon. Second, above and beyond these language-specific effects, the general impact of item length on NWR score may be relatively small, as shown in Piazzalunga et al.'s [-@piazzalunga2019articulatory] study on Italian and as borne out in the current dataset once controlling for other factors.

### 2021-01-20 manuscript.tex
started discussion

For consonants, we inspected complex ({[}ṭp{]}, {[}tp{]}, {[}kp{]}, {[}km{]}, {[}kṇ{]}, {[}mp{]}, {[}ɣ{]}, and {[}lβʲ{]}) versus simpler ones ({[}m{]}, {[}n{]}, {[}l{]}, {[}w{]}, {[}j{]}, {[}w{]}, {[}ṭ{]}, {[}g{]}, {[}p{]}, {[}t{]}, {[}k{]}, {[}f{]}, {[}h{]}, and {[}tʃ{]}), using the same logic: We looked at correct phone repetition, substitution with a change in complexity category, or a change within the same complexity category.\footnote{Note that the substitutions included phones that are not native to Yélî Dnye but do occur in English (e.g., {[}tʃ{]}); as these data come from careful transcriptions by a native Yélî Dnye speaker who is very fluent in English, we take these segmental substitutions faithfully as an indication that several of our participants have mastered production of some English phones, possibly produced within whole English word forms.}
The complex consonants in our stimuli occur in \textasciitilde{}17.33\% of languages' phonologies (range 0\% to 78\%);
whereas simple consonants in our stimuli occur in \textasciitilde{}67.62\% of languages' phonologies (range 13\% to 96\%).
Again these groups of sounds differ in their frequency within the language. Their type frequencies in Yélî Dnye are: complex consonants \textasciitilde{}0.04‰ (range 0.00‰ to 0.10‰) versus simple consonants \textasciitilde{}0.32‰ (range 0.06‰ to 0.55‰).

@@

Next, we inspected whether NWR scores varied as a function of word length (Table \ref{tab:tablength}). Participants scored much lower on monosyllables than on non-words of other lengths. This is likely due to the fact that the majority of monosyllables were designed to include sounds that are rare in the world's languages, which may indicate that they are hard to produce or perceive. Setting monosyllables aside, we observed the typical pattern of lower scores for longer items only for the whole-item scoring, and even there differences were rather small. In a generalized binomial mixed model excluding monosyllables, we included 479 observations, from 40 children producing, in any given trial, one of 24 potential target words. The analysis revealed a positive effect of age (ß = 0.56, SE ß = 0.14, p \textless{} 0.001) and a negative but non-significant estimate for target length in number of syllables (ß = -0.15, SE ß = 0.33, p = 0.65).

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Fig2-scores-by-age-1.pdf}
\caption{\label{fig:Fig2-scores-by-age}NWR whole-item scores for individual participants as a function of age and sex (purple = boys, orange = girls).}
\end{figure}

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure \ref{fig:Fig2-scores-by-age}, there was a greater deal of variance across the tested age range, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (5,649.08) = 0.47, p \textless{} 0.01). In contrast, there was no clear association between NWR scores and sex (Welch t (27.33) = -0.60, p = 0.56), birth order (data missing for 15 children, rho = (3,502.90) = -0.20, p = 0.33), or maternal education (rho ( 9,628.60) = 0.10, p = 0.55).

@@

We used non-word repetition to investigate phonological development in a language with a large phonological inventory (including some typologically rare segments). While the study, in itself, advances NWR research by demonstrating the successful application of this method in a rural, non-industrialized population (see also Cristia et al., 2020), the results also inform the larger body of NWR work regarding the influence of stimulus length, segmental frequency, age, and developmental context, where previous findings have often been mixed.

\hypertarget{item-complexity}{%
\subsubsection{Item complexity}\label{item-complexity}}

We investigated the effect of item complexity on NWR scores by varying both the number of syllables in the item and its average segmental frequency. Based on previous work, we had predicted that children would have higher NWR scores for shorter items. That said, previous work has shown both very small (Piazzalunga et al., 2019) and very large (Cristia et al., 2020; Jaber-Awida, 2018) effects of stimulus length and, further, the Yélî Dnye dictionary suggests that mono- and bi-syllabic words are nearly equally frequent in the current language, with trisyllabic and longer words making up a non-trivial 10\% of the remaining words. Compare this to, for example, English, which is substantially more skewed toward monosyllabic word forms \textbf{M2A: Alex I'm going off your note here (\enquote{Prediction for Yélî made before seeing the data: The length distribution in Yélî words is more balanced than that in English, and thus the score decline for poly- versus mono-syllables may be less pronounced than that for English.}"). I don't have a reference for this, can you please finish the thought or nix this bit?}. Setting aside our monosyllabic stimuli, which all contained typologically infrequent segments, we can examine effects of item length among the remaining stimuli, which range between 2 and 4 syllables long. While indeed NWR scores were overall lower for longer items (e.g., see Figure \ref{fig:prevlit-fig}), the effect of item length was not significant in a statistical model that additionally accounted for age and random effects of item and participant. In light of mixed prior results of item length, we propose two possible (and non-mutually exclusive) explanations for this minimal impact of item length. First, further extensions of this type of analysis in more populations may reveal that, in general (and cross-linguistically), item length effects are variable between languages, potentially reflecting the distribution of word lengths in the ambient language and other (morpho-)phonological tendencies in the lexicon. Second, above and beyond these language-specific effects, the general impact of item length on NWR score may be relatively small, as shown in Piazzalunga et al.'s (2019) study on Italian and as borne out in the current dataset once controlling for other factors.

Our monosyllabic items included typologically rare segments so that we could test whether lower average segmental frequency is associated with lower NWR scores. Typologically common sounds are associated with higher performance on a handful of other tasks (\textbf{REFS -- M2A: Alex, I added this based on your note, where it sounded like you had some particular studies in mind?}) though to our knowledge this has not yet been tested with non-word repetition. Regarding Yélî Dnye in particular, the phonemic inventory is both large and acoustically packed, in addition to containing several typologically infrequent (or unique) contrasts. We therefore expected to see that, while NWR scores would be lower for stimuli with lower average frequency, this effect would be relatively weak because the ambient language puts pressure on Yélî children to distinguish (perceptually and articulatorily) fine-grained phonetic differences in order to successfully communicate with others. Indeed we found a robust effect of average segmental frequency on NWR performance: Even accounting for age and random effects of item and participant, we see that target words with more frequent segments were repeated correctly more often. This effect is large, with a magnitude more than twice the size of the effect of participant age. This significant effect remains even once also accounting for the frequencies of these segments in Yélî Dnye child-directed speech, which are correlated with their typological frequencies. In sum, typological frequency effects, which have been found in other measurements of phonological processing. appear to strongly affect NWR performance, and do not appear mitigated by language-specific pressure to make finer-grained differences earlier in development.

@@

>>>>

For consonants, we inspected complex ({[}ṭp{]}, {[}tp{]}, {[}kp{]}, {[}km{]}, {[}kṇ{]}, {[}mp{]}, {[}ɣ{]}, and {[}lβʲ{]}) versus simpler ones ({[}m{]}, {[}n{]}, {[}l{]}, {[}w{]}, {[}j{]}, {[}w{]}, {[}ṭ{]}, {[}g{]}, {[}p{]}, {[}t{]}, {[}k{]}, {[}f{]}, {[}h{]}, and {[}tʃ{]}), using the same logic: We looked at correct phone repetition, substitution with a change in complexity category, or a change within the same complexity category.\footnote{Note that the substitutions included phones that are not native to Yélî Dnye but do occur in English (e.g., {[}tʃ{]}). These data come from careful transcriptions by a native Yélî Dnye speaker who is very fluent in English. This result suggests that several of our participants have mastered production of some English phones, possibly produced within whole English word forms.}
The complex consonants in our stimuli occur in \textasciitilde{}17.33\% of languages' phonologies (range 0\% to 78\%);
whereas simple consonants in our stimuli occur in \textasciitilde{}67.62\% of languages' phonologies (range 13\% to 96\%).
Again these groups of sounds differ in their frequency within the language. Their type frequencies in Yélî Dnye are: complex consonants \textasciitilde{}0.04‰ (range 0.00‰ to 0.10‰) versus simple consonants \textasciitilde{}0.32‰ (range 0.06‰ to 0.55‰).

@@

We next turned to our second research question by inspecting whether NWR scores varied as a function of word length (Table \ref{tab:tablength}). In this section and all subsequent ones, we only look at first attempts, for the reasons discussed previously. Additionally, we noticed that participants scored much lower on monosyllables than on non-words of other lengths. This is likely due to the fact that the majority of monosyllables were designed to include sounds that are rare in the world's languages, which may be harder to produce or perceive, as suggested by our previous analyses of NWR scores as a function of cross-linguistic phone frequency and error patterns. Therefore, we set monosyllables aside for this analysis.

We observed the typical pattern of lower scores for longer items only for the whole-item scoring, and even there differences were rather small. In a generalized binomial mixed model excluding monosyllables, we included 479 observations, from 40 children producing, in any given trial, one of 24 (non-monosyllabic) potential target words. The analysis revealed a positive effect of age (ß = 0.56, SE ß = 0.14, p \textless{} 0.001) and a negative but non-significant estimate for target length in number of syllables (ß = -0.15, SE ß = 0.33, p = 0.65).

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Fig2-scores-by-age-1.pdf}
\caption{\label{fig:Fig2-scores-by-age}NWR whole-item scores for individual participants as a function of age and sex (purple crosses = boys, orange circles = girls).}
\end{figure}

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals, as per our third research question. As shown in Figure \ref{fig:Fig2-scores-by-age}, there was a greater deal of variance across the tested age range, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (5,649.08) = 0.47, p \textless{} 0.01). In contrast, there was no clear association between NWR scores and sex (Welch t (27.33) = -0.60, p = 0.56), birth order (data missing for 15 children, rho = (3,502.90) = -0.20, p = 0.33), or maternal education (rho ( 9,628.60) = 0.10, p = 0.55).

@@

We used non-word repetition to investigate phonological development in a language with a large phonological inventory (including some typologically rare segments). We aimed to provide additional data on two questions already visitied in NWR work, namely the influence of stimulus length and individual variation, and one that has received less attention, regarding the possible relationship between phone frequency and NWR scores. An additional overarching goal was to revisit this task, which is very commonly used to document phonological development, particularly in children raised in urban settings, and who are learning IndoEuropean languages. We will thus consider our results not only in light of previous work, but also with attention to potential linguistic and population differences.

\hypertarget{associations-between-nwr-and-cross-linguistic-frequency}{%
\subsubsection{Associations between NWR and cross-linguistic frequency}\label{associations-between-nwr-and-cross-linguistic-frequency}}

Arguably the most innovative aspect of our data relate to the inclusion of phones that are less commonly found across languages, and rarely used in NWR tasks. Our monosyllabic items included typologically rare segments so that we could test whether lower average segmental frequency is associated with lower NWR scores. Typologically common sounds are associated with higher performance on a handful of other tasks (\textbf{REFS -- M2A: Alex, I added this based on your note, where it sounded like you had some particular studies in mind?}) though to our knowledge this has not yet been tested with non-word repetition. Regarding Yélî Dnye in particular, the phonemic inventory is both large and acoustically packed, in addition to containing several typologically infrequent (or unique) contrasts. We therefore expected to see that, while NWR scores would be lower for stimuli with lower average frequency, this effect would be relatively weak because the ambient language puts pressure on Yélî children to distinguish (perceptually and articulatorily) fine-grained phonetic differences in order to successfully communicate with others. Indeed we found a robust effect of average segmental frequency on NWR performance: Even accounting for age and random effects of item and participant, we see that target words with more frequent segments were repeated correctly more often. This effect is large, with a magnitude more than twice the size of the effect of participant age. This significant effect remains even once also accounting for the frequencies of these segments in Yélî Dnye child-directed speech, which are correlated with their typological frequencies. In sum, typological frequency effects, which have been found in other measurements of phonological processing. appear to strongly affect NWR performance, and do not appear mitigated by language-specific pressure to make finer-grained differences earlier in development.

@@

\hypertarget{item-length}{%
\subsubsection{Item Length}\label{item-length}}

We investigated the effect of item complexity on NWR scores by varying both the number of syllables in the item. Based on previous work, we had predicted that children would have higher NWR scores for shorter items. That said, previous work has shown both very small (Piazzalunga et al., 2019) and very large (Cristia et al., 2020; Jaber-Awida, 2018) effects of stimulus length and, further, the Yélî Dnye dictionary suggests that mono- and bi-syllabic words are nearly equally frequent in the current language, with trisyllabic and longer words making up a non-trivial 10\% of the remaining words. Compare this to, for example, English, which is substantially more skewed toward monosyllabic word forms \textbf{M2A: Alex I'm going off your note here (\enquote{Prediction for Yélî made before seeing the data: The length distribution in Yélî words is more balanced than that in English, and thus the score decline for poly- versus mono-syllables may be less pronounced than that for English.}"). I don't have a reference for this, can you please finish the thought or nix this bit?}. Setting aside our monosyllabic stimuli, which all contained typologically infrequent segments, we can examine effects of item length among the remaining stimuli, which range between 2 and 4 syllables long. While indeed NWR scores were overall lower for longer items (e.g., see Figure \ref{fig:prevlit-fig}), the effect of item length was not significant in a statistical model that additionally accounted for age and random effects of item and participant. In light of mixed prior results of item length, we propose two possible (and non-mutually exclusive) explanations for this minimal impact of item length. First, further extensions of this type of analysis in more populations may reveal that, in general (and cross-linguistically), item length effects are variable between languages, potentially reflecting the distribution of word lengths in the ambient language and other (morpho-)phonological tendencies in the lexicon. Second, above and beyond these language-specific effects, the general impact of item length on NWR score may be relatively small, as shown in Piazzalunga et al.'s (2019) study on Italian and as borne out in the current dataset once controlling for other factors.
